The \project{UNRAVE} catalog

### 2016-08-15 article/unrave.tex
Some stuff about data, etc

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\thecannon}{\project{The~Cannon}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\rave}{\project{\acronym{RAVE}}}
\newcommand{\logg}{\log g}
\newcommand{\teff}{T_{\mathrm{eff}}}
\newcommand{\argmin}[1]{\underset{#1}{\operatorname{argmin}}\,}

\newcommand{\Nstars}{483,330}

@@

% Chemistry and kinematics are essential for unfolding the Galaxy.

% Gaia will do this, and until Gaia releases abundances for stars from RVS spectra, we rely on ground-based surveys to provide chemical abundances.
% This includes GES, APOGEE, etc.

@@

The first \project{Gaia} data release includes positions, proper motions, and parallaxes for approximately two million stars in the Tycho-2 catalog \citep{tycho2}.  After cross-matching all major stellar spectroscopic surveys\footnote{Specifically we cross-matched the Tycho-2 catalog against {APOGEE} \citep{apogee}, \project{GALAH} \citep{galah}, \project{Gaia-ESO} \citep{ges}, \project{LAMOST} \citep{lamost}, and \project{RAVE} \citep{rave}}, we found the \project{RAVE} survey to have the largest overlap with the first \project{Gaia} data release: 292,036 stars.  We then used the \project{Gaia} universe model to estimate the precision in parallax and proper motions available in \project{Gaia} DR1 for every star in these cross-matches.  Comparing the expected precision to what is currently available, we further found that the \project{RAVE} survey would benefit most from \project{Gaia} DR1.  The distances of 63\% of the \project{RAVE}--\project{Gaia} DR1 overlap sample (182,862 stars) are expected to improve with the first \project{Gaia} data release, and 47\% are likely to have better proper motions (137,211 stars).

This motivated us to examine what chemical abundance information is currently available for \project{RAVE}, and to evaluate whether we could contribute additional abundances.

% We examined the overlap between different surveys and found RAVE to have the largest overlap with Gaia DR1.
% RAVE observed from X to Y, given blah blah blah resolution, etc
% RAVE was a radial velocity survey, so the S/N peaks at ~40
% Typically lower than the requisite S/N ratio required for chemical abundances.

% How many spectra, acquired over what time
% Data reduction
% Normalization
% Flux errors.

\section{Methods}
\label{sec:methods}

% Describe The Cannon briefly
% Initial tests using a single model for all RAVE spectra
% Complex structure led us to ivnestigate other models.
% Non-parametric form
% Censoring masks
% Priors on astrophysical parameters from isochrones

@@

% Chi-sq of the sample.
% Identify outliers.
% Stars with morphological classification

@@

% External validation: 
% --> RAVE Data Release 4
% --> Gaia-ESO Survey
% --> Asteroseismic nu_max and delta_nu available anywhere?
% --> Furhman and other studies in Kordopatis+ 2014

>>>>

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\thecannon}{\project{The~Cannon}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\rave}{\project{\acronym{RAVE}}}
\newcommand{\logg}{\log g}
\newcommand{\teff}{T_{\mathrm{eff}}}
\newcommand{\argmin}[1]{\underset{#1}{\operatorname{argmin}}\,}
\newcommand{\stub}[1]{\textbf{#1}}
\newcommand{\Nstars}{483,330}

@@

\stub{Chemistry and kinematics are essential for unfolding the Galaxy, etc.}

\stub{
Gaia will do this, and until Gaia releases abundances for stars from RVS 
spectra, we rely on ground-based surveys to provide chemical abundances.
This includes GES, APOGEE, etc.
}

The first \project{Gaia} data release includes positions, proper motions, and 
parallaxes for approximately two million stars in the Tycho-2 catalog \citep{tycho2}.  
After cross-matching all major stellar spectroscopic surveys\footnote{Specifically 
we cross-matched the Tycho-2 catalog against {APOGEE} \citep{apogee}, 
\project{GALAH} \citep{galah}, \project{Gaia-ESO} \citep{ges}, \project{LAMOST}
\citep{lamost}, and \project{RAVE} \citep{rave}}, we found the RAdial Velocity
Experiment (hereafter \rave) 
survey to have the largest overlap with the first \project{Gaia} data release: 
292,036 stars.  We then used the \project{Gaia} universe model \citep{GUMS} to 
estimate the precision in parallax and proper motions that will be available in 
\project{Gaia} DR1 for every star in these cross-matches.  Comparing the 
expected precision to what is currently available, we further found that the 
\project{RAVE} survey will benefit most from \project{Gaia} DR1.  The distances
of 63\% of the \project{RAVE}--\project{Gaia} DR1 overlap sample (182,862 stars)
are expected to improve with the first \project{Gaia} data release, and 47\% of
stars likely to have better proper motions (137,211 stars).  This motivated us to
examine what chemical abundance information is currently available for \rave, and
to evaluate whether we could contribute to this existing set.

We briefly describe the \rave\ data in Section \ref{sec:data}, before explaining
the model we employ in Section \ref{sec:model}.  In Section \ref{sec:validation}
we outline a number of validation experiments, including: internal sanity checks,
comparisons with literature samples, and investigations to ensure our results
are consistent with expectations from astrophysics.  We provide concluding
remarks in Section \ref{sec:conclusion}, as well as a description of how to
access the data.

@@

The \rave\ survey started observations using the 6dF system on the 1.2~m UK 
Schmidt telescope at the Australian Astronomical Observatory\footnote{Formerly 
the Anglo-Australian Observatory}.  The large 5.7~degree field-of-view and 
robotic fibre positioner made for very efficient observing: up to 150 targets 
could be simultaneously obtained.  When observations concluded in April 2013, a 
total of 574,630 spectra had been collected of 483,330 unique objects.

RAVE is a magnitude-limited survey, targeting stars with apparent $I$-band
magnitudes between $9 < I < 13$.  The $I$ band was used as the selection
magnitude because it overlaps with the wavelength range that \rave\ operates in:
8410--8795~\AA.  This setup includes the \ion{Ca}{2} near-infrared triplet,
strong stellar absorption lines that are visible even in very low S/N data 
(i.e., they are very useful for radial velocity determination).

% When did RAVE observe?
% Where from

% target selection.
% wavelength range

% How many data releases
% What was in previous data releases?

% Data reduction
% Normalization
% Flux errors

As the name implies, the \rave\ survey was principally designed to obtain radial
velocity measurements for up to $10^6$ stars.  Detailed chemical abundances were
a secondary science goal of the survey.  For this reason the distribution
of S/N ratios for \rave\ spectra is considerably lower than stellar 
spectroscopic surveys where chemical abundances are the defining objective.  The 
\rave\ spectra cover $\approx38$~nm in the infrared with an effective resolution
$\mathcal{R} \approx 7500$, and the S/N peaks at $\approx$50~per~pixel$^{-1}$.
For comparison, the \project{GALAH} survey \citep{DaSilva_2015} --- which was 
specifically constructed for detailed chemical abundance analyses --- includes a
wavelength range about 2.5 times larger at resolution $\mathcal{R} \approx 28,000$,
and target for S/N $\gtrsim100$.

Despite the relatively low resolution and S/N of the spectra, the \rave\ data
releases have demonstrated the team's enormous success in deriving radial 
velocities, stellar astrophysical parameters ($\teff$, $\logg$), as well as 
individual chemical abundances \citep{Steinmetz_2006,Zwitter_2008,Siebert_2011,
Kordopatis_2014, Kunder_2016}.  The data used for this analysis are the result
of a thorough reprocessing of all survey data.  This includes a detailed
re-reduction of the original data frames, an updated continuum-normalization
procedure, as well as an improvement on the quantification of flux errors in
each pixel.  At the end of this processing for each survey observation we are 
provided with: wavelengths, continuum-normalized fluxes, as well as 
$1\sigma$ uncertainties in flux values.  We refer the reader to the official 
fifth data release paper of the \rave\ survey, as presented by
\citet{Kunder_2016}, for more details of this reprocessing.

We chose not to re-normalize the spectra, as our tests proved to us that the
normalization procedure outlined in \citet{Kunder_2016} is sufficient for our
analysis.  We re-sampled all spectra onto a common rest-wavelength map.
Depending on the fibre used and stellar radial velocity, the rest-frame wavelength
range varied for each star.  Given that fluxes were unavailable in the edge
pixels for most stars, we excluded pixels outside of the rest wavelength range
$8423.2{\rm \AA} < \lambda < 8777.6{\rm \AA}$.

\section{Model}
\label{sec:model}

Given the relatively low S/N ratios in the \rave\ spectra, we chose to adopt a 
data-driven model (rather than a physics-driven model) for this analysis.

% Describe The Cannon briefly
% training set from APOGEE

% Initial tests using a single model for all RAVE spectra
% Censoring masks
% Priors on astrophysical parameters from isochrones

@@

% Chi-sq of the sample.
% Identify outliers.
% Convex hull?

% Stars with morphological classification

@@

% External validation: 
% --> RAVE Data Release 4
% --> Gaia-ESO Survey
% --> Asteroseismic nu_max and delta_nu available anywhere?
% --> Literature studies presented in Kordopatis+ 2014
